---
title: "Overlay Trait Data"
output: html_notebook
---

# Libraries

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(httr) # read out Dropbox folders
library(vegan)
library(readxl)
```

## Make species x trait matrix

Greenhouse traits

-   height, leaf dry matter content, specific leaf area, root mass fraction, relative growth rate, coarse root diameter, root density, specific root length of fine roots, specific root length of coarse roots

Field traits

-   height, leaf dry matter content, specific leaf area, seed mass, leaf C, leaf N

["H1:]{.underline} Pairing field results with a greenhouse 15 N experiment will allow us to relate species responses to their ability to uptake organic N"

"In addition, we will conduct a greenhouse experiment to isolate the mechanism (mineralized N vs organic N uptake rates ) by which compost may affect species composition"

Workflow plan:

-   overlay species trait data as vectors on the NMDS

    -   only greenhouse traits available (only sla data from field)

        -   use PC scores as composite trait axes OR

        -   average the traits for each species

## Read in and out matrices from traitOrdination code

```{r}
# write.csv(matrix.bray, "data/sitebyspecies_matrices/all_years_grazing_sitebyspecies_bray.csv", row.names = TRUE)
matrix.bray <- read.csv("data/sitebyspecies_matrices/all_years_grazing_sitebyspecies_bray.csv", row.names = 1)

# write.csv(matrix.block.bray, "data/sitebyspecies_matrices/all_years_grazing_sitebyspecies_bray_block.csv", row.names = TRUE)
matrix.block.bray <- read.csv("data/sitebyspecies_matrices/all_years_grazing_sitebyspecies_bray_block.csv", row.names = 1)


# write.csv(matrix.bray.low, "data/sitebyspecies_matrices/all_years_LOWgrazing_sitebyspecies_bray.csv", row.names = TRUE)
matrix.bray.low <- read.csv("data/sitebyspecies_matrices/all_years_LOWgrazing_sitebyspecies_bray.csv", row.names = 1)

# write.csv(matrix.bray.2021, "data/sitebyspecies_matrices/2021_grazing_sitebyspecies_bray.csv", row.names = TRUE)
matrix.bray.2021 <- read.csv("data/sitebyspecies_matrices/2021_grazing_sitebyspecies_bray.csv", row.names = 1)

# write.csv(matrix.bray.block.2021, "data/sitebyspecies_matrices/2021_grazing_sitebyspecies_bray_block.csv", row.names = TRUE)
matrix.bray.block.2021 <- read.csv("data/sitebyspecies_matrices/2021_grazing_sitebyspecies_bray_block.csv", row.names = 1)

# write.csv(matrix.bray.block.2021.low, "data/sitebyspecies_matrices/2021_LOWgrazing_sitebyspecies_bray_block.csv", row.names = TRUE)
matrix.bray.block.2021.low <- read.csv("data/sitebyspecies_matrices/2021_LOWgrazing_sitebyspecies_bray_block.csv", row.names = 1)

# write.csv(matrix.bray.block.2021.high, "data/sitebyspecies_matrices/2021_HIGHgrazing_sitebyspecies_bray_block.csv", row.names = TRUE)
matrix.bray.block.2021.high <- read.csv("data/sitebyspecies_matrices/2021_HIGHgrazing_sitebyspecies_bray_block.csv", row.names = 1)

# write.csv(matrix.bray.2019, "data/sitebyspecies_matrices/2019_grazing_sitebyspecies_bray.csv", row.names = TRUE)
matrix.bray.2019 <- read.csv("data/sitebyspecies_matrices/2019_grazing_sitebyspecies_bray.csv", row.names = 1)

# write.csv(matrix.bray.2020, "data/sitebyspecies_matrices/2020_grazing_sitebyspecies_bray.csv", row.names = TRUE)
matrix.bray.2020 <- read.csv("data/sitebyspecies_matrices/2020_grazing_sitebyspecies_bray.csv", row.names = 1)
```

### Read in trait data

```{r}
# # read in greenhouse trait data
# traits.gh <- read.csv("https://www.dropbox.com/scl/fi/c8vk31a807xg48jez1j03/GreenhouseTraits.csv?rlkey=ca736ea27k79mo7h8dbn67qiq&st=ya1ck6p7&dl=1")
# 
# # read in field trait data
# # download.file("https://www.dropbox.com/scl/fi/trqrnz54dkjoa9do4tle5/leaf-area-data_NG.xlsx?rlkey=s0kic2a5yqknikc4icek8y7rn&st=dl4x634h&dl=1",destfile = "field_sla.xlsx")
# traits.sla.field <- 
#   read_excel("field_sla.xlsx",col_names = TRUE,col_types = "text")
# 
# rda(traits.gh, scale = TRUE)
```

#### Investigate trait data, looking in Greenhouse trait folder

Trait data is identical. Will clean data in

```{r}
# dropbox.main <- read.csv("https://www.dropbox.com/scl/fi/c8vk31a807xg48jez1j03/GreenhouseTraits.csv?rlkey=ca736ea27k79mo7h8dbn67qiq&st=0h1bs53j&dl=1")
# 
# dropbox.gh <- read.csv("https://www.dropbox.com/scl/fi/67aaubdydiy4ixffmwvfe/GreenhouseTraits.csv?rlkey=rm2eqeqome7hgoedhnal0gx6d&st=vkym6evx&dl=1")
# 
# all.equal(dropbox.main, dropbox.gh)
# identical(dropbox.main, dropbox.gh)
```

```{r}
traits.cleaned <- read.csv("data/GreenhouseTraits_corrected_20240212.csv") %>%
  mutate(
    Fresh.leaf.mass..g. = as.numeric(ifelse(
      Fresh.leaf.mass..g. == "Skipped accidentally", NA, Fresh.leaf.mass..g.
    )),
    Dry.leaf.mass..g. = as.numeric(ifelse(
      Dry.leaf.mass..g. == "No sample", NA, Dry.leaf.mass..g.
    )),
    Root.dry.biomass..g. = as.numeric(ifelse(
      Root.dry.biomass..g. == "roots lost", NA, Root.dry.biomass..g.
    )),
    Root.volume..cm3. = as.numeric(ifelse(
      Root.volume..cm3. == "not scanned", NA, Root.volume..cm3.
    ))
  ) 
```

### Summarize trait data (GH) for each species

#### means

```{r}
# make trait species ID match matrix data (notes shown below) -------
traits.cleaned <- traits.cleaned %>%
  mutate(
    ID = ifelse(
      ID == "Agoseris", "AGSP", ifelse(
        ID == "AVEBAR", "AVBA", ifelse(
          ID == "BRDR", "BRDI", ID
        )
      )
    )
  )

# summary(traits.cleaned)
traits.means <- traits.cleaned %>%
  group_by(ID) %>%
  summarise(
    across(4:(ncol(traits.gh) - 1), list(mean = ~mean(.x, na.rm = TRUE)
  ))) %>%
  rename(
    Height = Height..cm._mean,
    FreshLeafMass = Fresh.leaf.mass..g._mean,
    DryLeafMass = Dry.leaf.mass..g._mean,
    LDMC = LDMC_mean, # Leaf dry matter content (LDMC, the ratio of leaf dry mass to fresh mass)
    LeafArea = Leaf.Area..cm2._mean,
    SLA = SLA..cm2.g._mean,
    ShootDryBiomass = Shoot.dry.biomass..g._mean,
    RootDryBiomass = Root.dry.biomass..g._mean,
    TotalBiomass = Total.biomass..g._mean,
    RMF = RMF_mean, # Root mass fraction (RMF) is a plant trait that measures the proportion of a plant's dry mass that is in its roots
    RootVol = Root.volume..cm3._mean,
    RootDensity = Root.density..g.cm3._mean,
    CoarseRootDiameter = Coarse.root.diameter..mm._mean,
    RootLength = Length..mm._mean,
    FineRootLength = Fine.root.length..mm._mean,
    CoarseRootLength = Coarse.root.length..mm._mean,
    CoarseRootSpecLength = Coarse.root.specific.length..cm.g._mean,
    FineRootSpecLength = Fine.root.specific.length..cm.g._mean,
    PropFineRoots = Proportion.fine.roots_mean
  )
  

# make site by species matrix match traits species
matrix.names <- data.frame(colnames(matrix.bray)) #species comp (75 count)
trait.names <- data.frame(traits.means$ID) #trait species (80 count)

intersect.names <- intersect(colnames(matrix.bray), traits.means$ID) # same species (43 count)
setdiff(colnames(matrix.bray), traits.means$ID)
setdiff(traits.means$ID, colnames(matrix.bray))

# check matrix codes to see if same species are named different things
unique.matrix.species <- cover.dat %>%
  dplyr::select(code4, species) %>%
  summarise(
    code4 = unique(code4),
    species = unique(species)
  )

unique.trait.species <- traits.cleaned %>%
  dplyr::select(Taxon, ID) %>%
  distinct(ID, .keep_all = TRUE)
  

# write traits.means csv ----
# write.csv(traits.means, "data/traits/all_sp_trait_means.csv")
 

# write.csv(traits.means.comp, "data/traits/field_sp_trait_means.csv")
```

### <!--#  -->See if I have to match names with traits ever year (ANS: All Same Species)

```{r}
# # all years -------------
# # make site by species matrix match traits species
# matrix.names <- data.frame(colnames(matrix.bray)) #species comp (75 count)
# trait.names <- data.frame(traits.means$ID) #trait species (80 count)
# 
# intersect.names <- intersect(colnames(matrix.bray), traits.means$ID) # same species (43 count)
# setdiff(colnames(matrix.bray), traits.means$ID)
# setdiff(traits.means$ID, colnames(matrix.bray))
# 
# # check matrix codes to see if same species are named different things
# unique.matrix.species <- cover.dat %>%
#   dplyr::select(code4, species) %>%
#   summarise(
#     code4 = unique(code4),
#     species = unique(species)
#   )
# 
# unique.trait.species <- traits.cleaned %>%
#   dplyr::select(Taxon, ID) %>%
#   distinct(ID, .keep_all = TRUE)
#   
# # 2019 -------------
# data.frame(colnames(matrix.bray.2019)) #species comp (75 count)
# data.frame(colnames(matrix.bray.2020)) #species comp (75 count)
# data.frame(colnames(matrix.bray.2021)) #species comp (75 count)
```

-   ***change*** Trait name: 'Agoseris' --\> 'Agoseris unknown'; assuming the same as 'AGSP' --\> 'Agoseris sp.' in matrix data (change to matrix name)

-   [What's the difference between:]{.underline}

    -   'ANAR' and 'ANARp' in trait data?

    -   'BRHO' and 'BRHOp' in trait data?

    -   'BRNIf' and 'BRNIp' in trait data? (not in matrix data)

    -   'CLPUf' and 'CLPUp' in trait data? (not in matrix data)

    -   'GITRf' and 'GITRp' in trait data? (not in matrix data)

    -   'LENIf' and 'LENIp' in trait data? (not in matrix data)

    -   'TRHI' and 'TRHIpi' (same sp name) in trait data? (TRHI in matrix data)

    -   'TRWIf' and 'TRWIp' and 'TRWIpi' in trait data? (not in matrix data)

    -   'Crassula' and 'Crassula tillaea' (could they be the same?)

    -   'Sanicula' (unknown) and 'SABI1' or 'SABI2' (could they be the same?)

    -   'MAELf' (trait data) and 'MAD1' (Madia sp. 1 ("tall tarweed")' (could they be the same?)

    -   'MAELp' (trait data) and 'MAD2' (Madia sp. 1 ("small tarweed")' (could they be the same?)

    -   'Trifolium eriocephalum' (TRER) and 'Triphysaria eriantha' (TRER) (could they be the same?)

-   ***change*** 'AVEBAR' (trait data) to 'AVBA' (matrix data)

-   ***change*** BRDR --\> BRDI

## Overlay Trait Vectors onto NMDS

[**Aboveground Traits**]{.underline}

-   Height

-   **Leaf**

    -   FreshLeafMass

    -   DryLeafMass

    -   LDMC

        -   Leaf dry matter content (LDMC, the ratio of leaf dry mass to fresh mass)

    -   LeafArea

    -   SLA

    -   ShootDryBiomass

[**Belowground Traits**]{.underline}

-   RootDryBiomass

-   RMF

    -   Root mass fraction (RMF) is a plant trait that measures the proportion of a plant's dry mass that is in its roots

-   RootVol

-   RootDensity

-   RootLength

-   PropFineRoots

-   **Coarse Roots**

    -   CoarseRootDiameter

    -   CoarseRootLength

    -   CoarseRootSpecLength

-   **Fine Roots**

    -   FineRootLength

    -   FineRootSpecLength

-   TotalBiomass

### ALL YEARS -- LOW AND HIGH GRAZING

#### Prepare NMDS and Traits

Stress:

```         
0.2011127
```

```{r}
# subset nmds to match species names
nmds.comp.trait <- metaMDS(matrix.bray %>%
                             dplyr::select(intersect.names), 
                           k=2, trymax = 25)
nmds.comp.trait # pretty not great stress
stressplot(nmds.comp.trait)
plot(nmds.comp.trait, type="t")

nmds1 <- as.data.frame(scores(nmds.comp.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.comp.trait, choices=c(2), display=c("sites")))

nmds_dat_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# subset matrix and trait data to match names
traits.means.comp <- traits.means %>%
  filter(ID %in% intersect.names) 

# traits.means.comp2 <- traits.means.comp %>%
# dplyr::select(-ID)
rownames(traits.means.comp2) <- traits.means.comp$ID

# trait vectors onto the NMDS
# envfit(nmds.comp.trait, traits.means.comp, permutations = 999)
# example from stack overflow: envfit(ord ~ var1 + var2 + var3, dune.spec, display="sp"), 'ord' is the nmds model, 'var1' is the trait1 i think?

colnames(traits.means.comp)
vectors.nmds.comp.trait <- envfit(nmds.comp.trait ~ Height + 
                                    LeafArea +
                                    RMF, traits.means.comp, display = "sp") # ONLY Height, Leaf area, and RMF significant

#TEMPLATE: 
  # envfit(nmds.block.trait ~ Height + 
  #          FreshLeafMass +
  #          DryLeafMass +
  #          LDMC +
  #          LeafArea +
  #          SLA +
  #          ShootDryBiomass +
  #          RootDryBiomass +
  #          TotalBiomass +
  #          RMF +
  #          RootVol +
  #          RootDensity +
  #          CoarseRootDiameter +
  #          RootLength +
  #          FineRootLength +
  #          CoarseRootLength +
  #          CoarseRootSpecLength +
  #          FineRootSpecLength +
  #          PropFineRoots, traits.means.comp, display = "sp", na.rm = TRUE)

# is it different when you only do certain traits (above- vs. below-ground)? or can you just throw everything in there and it will tell you what is significant and you can subset from there?

# Looks like different subsets of traits put in are the same, so I shall proceed with the mega models and verify with Lauren

trait_vectors <- as.data.frame(vectors.nmds.comp.trait$vectors$arrows)

```

NMDS model: only Height, Leaf area, and RMF () found significant.

#### Plot

```{r}
ggplot(nmds_dat_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  labs(title = "Grouped by Nutrient Treatment") +
  custom_colors +
  scale_color_manual(
    values = c(
      "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
      "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
      "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
      "wet" = rgb(0, 0, 255, maxColorValue = 255),
      "dry" = rgb(183, 65, 14, maxColorValue = 255)
    )
  ) +
  geom_segment(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = NMDS1, y = NMDS2, label = rownames(vectors.nmds.comp.trait$vectors$arrows)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_trait, aes(NMDS1, NMDS2, color = as.factor(yr), fill = as.factor(yr))) +
  geom_point(shape = 21, size = 4, stroke = 0.5) +
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  labs(title = "Grouped by Year")+
  geom_segment(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = NMDS1, y = NMDS2, label = rownames(vectors.nmds.comp.trait$vectors$arrows)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_trait, aes(NMDS1, NMDS2, fill = as.factor(ppt_trt), color = as.factor(ppt_trt))) +
  geom_point(shape = 21, size = 4, stroke = 0.5) +
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  labs(title = "Grouped by Precipitation Treatment") +
  custom_colors +
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
)+
  geom_segment(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = NMDS1, y = NMDS2, label = rownames(vectors.nmds.comp.trait$vectors$arrows)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_trait, aes(NMDS1, NMDS2,  fill = as.factor(grazing_hist), color = as.factor(grazing_hist))) +
  geom_point(shape = 21, size = 4, stroke = 0.5) +
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  labs(title = "Grouped by Grazing History") +
  geom_segment(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
    aes(x = NMDS1, y = NMDS2, label = rownames(vectors.nmds.comp.trait$vectors$arrows)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

# ggplot(nmds_dat_trait, aes(NMDS1, NMDS2,  color = as.factor(grazing_hist), fill = as.factor(yr))) +
#   geom_point(aes(fill = as.factor(yr)), shape = 21, size = 4, stroke = 1) +
#   stat_ellipse(aes(color = as.factor(yr)),
#                 geom = "polygon",
#                fill = NA,
#                size = 1) +
#   labs(title = "Grouped by Grazing History & Year") +
#   scale_color_manual(
#   values = c(
#     "low" = "cyan",
#     "high" = "brown",
#   "2019" = "salmon",
#   "2020" = "limegreen",
#   "2021" = "cornflowerblue")
#   ) +
#   geom_segment(
#     data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
#     aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
#     inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
#     arrow = arrow(length = unit(0.2, "cm")), 
#     color = "red", 
#     size = 1
#   ) +
#   geom_text(
#     data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
#     aes(x = NMDS1, y = NMDS2, label = rownames(vectors.nmds.comp.trait$vectors$arrows)), 
#     inherit.aes = FALSE,  # Avoids unwanted inheritance
#     hjust = -0.2, 
#     vjust = -0.2, 
#     color = "red",
#     size = 5
#   )
#   
# ggplot(nmds_dat_trait, aes(NMDS1, NMDS2, color = nut_trt, fill = as.factor(yr))) +
#   geom_point(shape = 21, size = 4, stroke = 1) +
#   stat_ellipse(aes(color = as.factor(nut_trt)),
#              geom = "polygon",
#              fill = NA,  # Makes the ellipse transparent inside
#              size = 1) +
#   labs(title = "Grouped by Nutrient Treatment & Year") +
#   scale_color_manual(
#   values = c(
#     "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
#     "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
#     "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
#     "wet" = rgb(0, 0, 255, maxColorValue = 255),
#     "dry" = rgb(183, 65, 14, maxColorValue = 255)
#   )
# ) +
#   geom_segment(
#     data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
#     aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
#     inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
#     arrow = arrow(length = unit(0.2, "cm")), 
#     color = "red", 
#     size = 1
#   ) +
#   geom_text(
#     data = as.data.frame(vectors.nmds.comp.trait$vectors$arrows), 
#     aes(x = NMDS1, y = NMDS2, label = rownames(vectors.nmds.comp.trait$vectors$arrows)), 
#     inherit.aes = FALSE,  # Avoids unwanted inheritance
#     hjust = -0.2, 
#     vjust = -0.2, 
#     color = "red",
#     size = 5
#   )
```

### ALL YEARS -- LOW AND HIGH GRAZING (Blocks)

Stress:

```         
0.2353937
```

#### Prepare NMDS and Traits

```{r}
# subset nmds to match species names
nmds.block.trait <- metaMDS(matrix.block.bray %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.block.trait # pretty not great stress

nmds1 <- as.data.frame(scores(nmds.block.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.block.trait, choices=c(2), display=c("sites")))

nmds_dat_block_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data
traits.means.comp 

# trait vectors onto the NMDS
vectors.nmds.block.trait <-
  envfit(nmds.block.trait ~ Height + 
           DryLeafMass +
           LeafArea +
           RootDryBiomass +
           RMF +
           RootDensity, traits.means.comp, display = "sp")
# ONLY Height, DryLeafMass, LeafArea, RMF significant (RootDryBiomass, RootDensity marginal)

trait_vectors <- as.data.frame(vectors.nmds.block.trait$vectors$arrows)

```

#### Plot

```{r}
ggplot(nmds_dat_block_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  labs(title = "Grouped by Nutrient Treatment - Blocks w Traits") +
  custom_colors +
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_block_trait, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  labs(title = "Grouped by Precipitation Treatment - Blocks w Traits") +
  custom_colors +
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )




```

### **All years - LOW GRAZING ONLY**

#### Prepare NMDS and Traits

Stress:

```         
0.1879671
```

```{r}
# subset nmds to match species names
nmds.low.trait <- metaMDS(matrix.bray.low %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.low.trait 

nmds1 <- as.data.frame(scores(nmds.low.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.low.trait, choices=c(2), display=c("sites")))

nmds_dat_low_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data: traits.means.comp 

# trait vectors onto the NMDS
vectors.nmds.low.trait <-
  envfit(nmds.low.trait ~ FreshLeafMass +
           RMF, traits.means.comp, display = "sp", na.rm = TRUE)
# ONLY FreshLeafMass (RMF marginal)

trait_vectors <- as.data.frame(vectors.nmds.low.trait$vectors$arrows)

```

#### Plot

```{r}
ggplot(nmds_dat_low_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  labs(title = "2019-2021 (LOW) w Traits - Grouping by Nutrient Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  custom_colors + 
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_low_trait, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  labs(title = "2019-2021 (LOW) w Traits - Grouping by Precipitation Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  custom_colors + 
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )
```

### 2019

Stress:

```         
0.06290431
```

#### Prepare NMDS and Traits

```{r}
# subset nmds to match species names
nmds.2019.trait <- metaMDS(matrix.bray.2019 %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.2019.trait 

nmds1 <- as.data.frame(scores(nmds.2019.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.2019.trait, choices=c(2), display=c("sites")))

nmds_dat_2019_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data: traits.means.comp 


# trait vectors onto the NMDS
vectors.nmds.2019.trait <-
  envfit(nmds.2019.trait ~ Height + 
           DryLeafMass +
           LeafArea +
           RMF +
           RootDensity +
           CoarseRootDiameter, traits.means.comp, display = "sp", na.rm = TRUE)

vectors.nmds.2019.trait
# ONLY Height, LeafArea (DryLeafMass, RMF, RootDensity, CoarseRootDiameter marginal)

trait_vectors <- as.data.frame(vectors.nmds.2019.trait$vectors$arrows)
```

#### Plot

```{r}
ggplot(nmds_dat_2019_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2019 w Traits - Grouping by Nutrient Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_2019_trait, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2019 w Traits - Grouping by Precipitation Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )
```

### 2020

Stress:

```         
0.1512864
```

#### Prepare NMDS and Traits

```{r}
# subset nmds to match species names
nmds.2020.trait <- metaMDS(matrix.bray.2020 %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.2020.trait 

nmds1 <- as.data.frame(scores(nmds.2020.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.2020.trait, choices=c(2), display=c("sites")))

nmds_dat_2020_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data: traits.means.comp 


# trait vectors onto the NMDS
vectors.nmds.2020.trait <-
  envfit(nmds.2020.trait ~ FreshLeafMass +
           LeafArea, traits.means.comp, display = "sp", na.rm = TRUE)

# ONLY FreshLeafMass, LeafArea

trait_vectors <- as.data.frame(vectors.nmds.2020.trait$vectors$arrows)
```

#### Plot

```{r}
ggplot(nmds_dat_2020_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2020 w Traits - Grouping by Nutrient Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_2020_trait, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2020 w Traits - Grouping by Precipitation Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )
```

### 2021

#### Prepare NMDS and Traits

Stress:

```         
0.123061
```

```{r}
# subset nmds to match species names
nmds.2021.trait <- metaMDS(matrix.bray.2021 %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.2021.trait 

nmds1 <- as.data.frame(scores(nmds.2021.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.2021.trait, choices=c(2), display=c("sites")))

nmds_dat_2021_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data: traits.means.comp 
rownames(traits.means.comp) <- traits.means.comp$ID


# trait vectors onto the NMDS
vectors.nmds.2021.trait <-
  envfit(nmds.2021.trait ~ Height + 
           FreshLeafMass +
           LeafArea, na.omit(traits.means.comp), display = "sp", na.rm = TRUE)

# (Height, FreshLeafMass, LeafArea marginal)

trait_vectors <- as.data.frame(vectors.nmds.2021.trait$vectors$arrows)
```

#### Plot

```{r}
ggplot(nmds_dat_2021_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 w Traits - Grouping by Nutrient Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_2021_trait, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 w Traits - Grouping by Precipitation Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )
```

### 2021 -- Blocks

Stress:

```         
0.170799
```

#### Prepare NMDS and Traits

```{r}
# subset nmds to match species names
nmds.2021.block.trait <- metaMDS(matrix.bray.block.2021 %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.2021.block.trait 

nmds1 <- as.data.frame(scores(nmds.2021.block.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.2021.block.trait, choices=c(2), display=c("sites")))

nmds_dat_2021_block_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data: traits.means.comp 


# trait vectors onto the NMDS
vectors.nmds.2021.block.trait <-
  envfit(nmds.2021.block.trait ~ DryLeafMass +
           LeafArea, traits.means.comp, display = "sp", na.rm = TRUE)

# ONLY DryLeafMass, LeafArea

trait_vectors <- as.data.frame(vectors.nmds.2021.block.trait$vectors$arrows)

```

#### Plot

```{r}
ggplot(nmds_dat_2021_block_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 (Block) w Traits - Grouping by Nutrient Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_2021_block_trait, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 (Block) w Traits - Grouping by Precipitation Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )
```

### 2021 -- Blocks (LOW)

Stress:

```         
0.1299711
```

#### Prepare NMDS and Traits

```{r}
# subset nmds to match species names
nmds.2021.block.low.trait <- metaMDS(matrix.bray.block.2021.low %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.2021.block.low.trait 

nmds1 <- as.data.frame(scores(nmds.2021.block.low.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.2021.block.low.trait, choices=c(2), display=c("sites")))

nmds_dat_2021_block_low_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data: traits.means.comp 


# trait vectors onto the NMDS
vectors.nmds.2021.block.low.trait <-
  envfit(nmds.2021.block.low.trait ~ FreshLeafMass, traits.means.comp, display = "sp", na.rm = TRUE)

# ONLY FreshLeafMass

trait_vectors <- as.data.frame(vectors.nmds.2021.block.low.trait$vectors$arrows)
```

#### Plot

```{r}
ggplot(nmds_dat_2021_block_low_trait, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 (Block; LOW) w Trait - Grouping by Nutrient Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )

ggplot(nmds_dat_2021_block_low_trait, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 (Block; LOW) w Trait - Grouping by Precipitation Treatment") + 
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
) +
  geom_segment(
    data = trait_vectors, 
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
    inherit.aes = FALSE,  # 🚀 Prevents it from looking for 'nut_trt'
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red", 
    size = 1
  ) +
  geom_text(
    data = trait_vectors, 
    aes(x = NMDS1, y = NMDS2, label = rownames(trait_vectors)), 
    inherit.aes = FALSE,  # Avoids unwanted inheritance
    hjust = -0.2, 
    vjust = -0.2, 
    color = "red",
    size = 5
  )
```

### 2021 -- Blocks (HIGH)

#### Prepare NMDS and Traits

```{r}
# subset nmds to match species names
nmds.2021.block.high.trait <- metaMDS(matrix.bray.block.2021.high %>%
                             dplyr::select(intersect.names), , k=2, trymax = 25)
nmds.2021.block.high.trait 

nmds1 <- as.data.frame(scores(nmds.2021.block.high.trait, choices=c(1), display=c("sites")))
nmds2 <- as.data.frame(scores(nmds.2021.block.high.trait, choices=c(2), display=c("sites")))

nmds_dat_2021_block_high_trait <- cbind(nmds1, nmds2) %>%
  as.data.frame() %>%  # Ensure it's a data frame
  rownames_to_column(var = "site_id") %>%  # Move row names into a new column
  separate(site_id, into = c("nut_trt", "ppt_trt", "yr", "grazing_hist"), sep = "_")


# trait data: traits.means.comp 


# trait vectors onto the NMDS
vectors.nmds.2021.block.high.trait <-
  envfit(nmds.2021.block.high.trait ~ Height + 
           FreshLeafMass +
           DryLeafMass +
           LDMC +
           LeafArea +
           SLA +
           ShootDryBiomass +
           RootDryBiomass +
           TotalBiomass +
           RMF +
           RootVol +
           RootDensity +
           CoarseRootDiameter +
           RootLength +
           FineRootLength +
           CoarseRootLength +
           CoarseRootSpecLength +
           FineRootSpecLength +
           PropFineRoots, traits.means.comp, display = "sp", na.rm = TRUE)

# None

trait_vectors <- as.data.frame(vectors.nmds.2021.block.high.trait$vectors$arrows)
```

#### Plot

```{r}
ggplot(nmds_2021_dat.block.high, aes(NMDS1, NMDS2,  fill = nut_trt, color = nut_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 (Block; HIGH) - Grouping by Nutrient Treatment") +
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
)

ggplot(nmds_2021_dat.block.high, aes(NMDS1, NMDS2,  fill = ppt_trt, color = ppt_trt)) +
  geom_point(shape = 21, size = 4, stroke = 0.5) + 
  custom_colors + 
  labs(title = "2021 (Block; HIGH) - Grouping by Precipitation Treatment") +
  stat_ellipse(type = "norm", size = 1, alpha = 0.5) + # Ellipses for each group
  scale_color_manual(
  values = c(
    "control" = rgb(220, 220, 220, maxColorValue = 255),      # Lighter Gray
    "+N fertilizer" = rgb(140, 40, 150, maxColorValue = 255), # Vibrant Purple
    "+compost" = rgb(220, 180, 60, maxColorValue = 255),       # Bright Golden Yellow
    "wet" = rgb(0, 0, 255, maxColorValue = 255),
    "dry" = rgb(183, 65, 14, maxColorValue = 255)
  )
)
```
