# Exploratory Plant Community Analysis

1.  Species responses

    1.  **individual species**

    2.  natives vs. exotics

    3.  life forms

    4.  **between years**

    5.  between treatments

    6.  between competitive groups

    7.  traits

        1.  above- and belowground

[*Initial Goal:*]{.underline} *Get a sense of the general winners and losers of each treatments, then relate them by traits in ordinational space.*

Author: Caryn D. Iwanaga

Updated: 12/12/2024

## Libraries

```{r library, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(httr) # read out Dropbox folders
```

## Read in Data

```{r readDat}

# cover data -----
cover.dat <- read.csv("https://www.dropbox.com/scl/fi/up8nnzkcpchsr45f8cm92/Compost_Cover_LongClean.csv?rlkey=z2tvaj8t6khadef7ydz782zka&st=qwef9ys0&dl=1")

cover.wide <- read.csv("https://www.dropbox.com/scl/fi/lzgpi7ofpkrf5fk1d2e73/Compost_Cover_WideClean.csv?rlkey=7vzoz0u88esgvaz2s3rnmcx29&st=nwdfgtyu&dl=1")

# treatment data ----
trt.codes <- read.csv("https://www.dropbox.com/scl/fi/czb2oehii9m1n0iill57k/Compost_TreatmentKey.csv?rlkey=r9yu1c89o76lljqpe91g05uoi&st=lixbigdu&dl=1")

sp.list <- read.csv("https://www.dropbox.com/scl/fi/b43mka47sp7hfubz2vsi1/Compost_SppList.csv?rlkey=auz8vb9ftjlfdtc4q64f8m9ax&st=3p7hex2z&dl=1")
```

## Individual Species Responses (exploratory)

Divide the individual species responses to treatments, year, plots and see how they do

Could group species and see if there's anything there as well.

4 blocks

36 plots

2019 - 2020 years

+-------------------------------------+--------------------------+-------------+----------------+
| Nutrient Treatments (confirm codes) | Precipitation Treatments | Year        | Sample Events  |
+=====================================+==========================+=============+================+
| Compost (C)                         | Dry (D)                  | 2019        | 1 (2019-04-19) |
|                                     |                          |             |                |
|                                     |                          |             | 2 (2019-05-07) |
+-------------------------------------+--------------------------+-------------+----------------+
| N Fertilizer (F)                    | Control (XC)             | 2020        | 1 (2020-04-28) |
|                                     |                          |             |                |
|                                     |                          |             | 2 (NA)         |
+-------------------------------------+--------------------------+-------------+----------------+
| Control or None (N)                 | Wet (W)                  | 2021        | 1 (2021-04-07) |
|                                     |                          |             |                |
|                                     |                          |             | 2 (2021-05-13) |
+-------------------------------------+--------------------------+-------------+----------------+

```{r}
cover.dat %>%
  group_by(fxnl_grp, nativity, genus, epithet, code6) %>%
  summarise(unique(species))
  

```

Starting species:

1.  Avena barbata
2.  Taeniatherum caput-medusae
3.  Hordeum murinum
4.  Aira caryophyllea
5.  AST1
6.  Bromus diandrus

make boxplots of cover values of species with different treatments, years, prec.

```{r Avena}
response.avebar <- cover.dat %>%
  filter(code6 == "AVEBAR") %>%
  group_by(nut_trt, ppt_trt, yr) %>%
  mutate( 
    nut_trt = factor(ifelse(nut_trt =="C", "+compost", ifelse(nut_trt =="F", "+N fertilizer", ifelse(nut_trt =="N", "control", NA))), levels = c("control", "+N fertilizer", "+compost")),
    ppt_trt = ifelse(ppt_trt == "D", "dry", ifelse(ppt_trt == "XC", "control", ifelse(ppt_trt == "W", "wet", NA)))) %>%
  summarise(cover = mean(pct_cover), sd = sd(pct_cover)) %>%
  ggplot(aes(nut_trt, cover, fill = nut_trt)) +
  geom_col(position = "dodge") +
  geom_col(position = position_dodge(width = 0.9), width = 0.7) +  # Bar plot
  geom_errorbar(
    aes(ymin = cover - sd, ymax = cover + sd),  # Error bars
    position = position_dodge(width = 0.9),  # Align with bars
    width = 0.4,
    alpha = 0.3
  ) +
   scale_fill_manual(
    values = c(
      "control" = rgb(195, 197, 193, maxColorValue = 250),
      "+N fertilizer" = rgb(87, 62, 92, maxColorValue = 225),
      "+compost" = rgb(167, 156, 109, maxColorValue = 225)
    )
  ) +
  facet_grid(yr~ppt_trt) +
  theme_bw()+
  labs(
    title = "A.barbata cover response (2019-2021)"
  )

# ggsave(
#   filename = "AVEBAR_cover_TrtsYr.png",
#   plot = response.avebar,
#   width = 11,
#   height = 7)
```

```{r looksee}
cover.dat %>%
  filter(code6 == "AVEBAR") %>%
  group_by(nut_trt, ppt_trt, yr) %>%
  # group_by(nut_trt, yr) %>%
  mutate( 
    nut_trt = factor(ifelse(nut_trt =="C", "+compost", ifelse(nut_trt =="F", "+N fertilizer", ifelse(nut_trt =="N", "control", NA))), levels = c("control", "+N fertilizer", "+compost")),
    ppt_trt = ifelse(ppt_trt == "D", "dry", ifelse(ppt_trt == "XC", "control", ifelse(ppt_trt == "W", "wet", NA)))) %>%
  filter(fulltrt == "NXC", yr == 2019)
```
