```{r}
library(dplyr)
library(vegan)
library(tidyr)
library(ggplot2)
```

```{r}
# read in greenhouse trait data
traits.gh <- read.csv("data/GreenhouseTraits.csv")
```

Cleaning:

-   make numeric class (

    ```         
    Fresh.leaf.mass..g. Dry.leaf.mass..g.Root.dry.biomass..g.Root.volume..cm3.
    ```

```{r}
traits.gh.cleaned <- traits.gh %>%
  mutate(
    Fresh.leaf.mass..g. = as.numeric(ifelse(
      Fresh.leaf.mass..g. == "Skipped accidentally", NA, Fresh.leaf.mass..g.
    )),
    Dry.leaf.mass..g. = as.numeric(ifelse(
      Dry.leaf.mass..g. == "No sample", NA, Dry.leaf.mass..g.
    )),
    Root.dry.biomass..g. = as.numeric(ifelse(
      Root.dry.biomass..g. == "roots lost", NA, Root.dry.biomass..g.
    )),
    Root.volume..cm3. = as.numeric(ifelse(
      Root.volume..cm3. == "not scanned", NA, Root.volume..cm3.
    ))
  ) 

traits.gh.summary <- traits.gh.cleaned %>%
  group_by(Taxon) %>%
  summarise(across(
    .cols = where(is.numeric),  # Apply to all numeric columns
    .fns = list(
      min = ~ min(., na.rm = TRUE),
      max = ~ max(., na.rm = TRUE),
      mean = ~ mean(., na.rm = TRUE),
      median = ~ median(., na.rm = TRUE),
      q1 = ~ quantile(., 0.25, na.rm = TRUE),
      q3 = ~ quantile(., 0.75, na.rm = TRUE)
    ),
    .names = "{.col}_{.fn}"  # Custom column names
  )) %>%
  pivot_longer(
    cols = c(Height..cm._min:Proportion.fine.roots_q3)
  ) %>%
  pivot_wider(
    names_from = Taxon,
    values_from = value
  ) %>%
  mutate(
    round(traits.gh.summary[,c(2:71)],4),
    rowname = name
  )

# write.csv(traits.gh.summary, "data/inspect_ghdat.csv", row.names = FALSE)

summary(traits.gh.cleaned)

hist(traits.gh$Height..cm.)
summary(traits.gh)

inspect.sd <- traits.gh.cleaned %>%
  group_by(Taxon) %>%
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = list(
        mean = ~ mean(., na.rm = TRUE),
        sd = ~ sd(., na.rm = TRUE)
      ),
    .names = "{.col}_{.fn}"  # Custom column names
    )
  )

# write.csv(inspect.sd, "data/inspectSD_ghdat.csv", row.names = FALSE)
```

Inspect ranges to find any misentered data

[Suspicious Data checking:]{.underline}

<table>
<colgroup>
<col style="width: 29%" />
<col style="width: 21%" />
<col style="width: 21%" />
<col style="width: 17%" />
<col style="width: 16%" />
<col style="width: 30%" />
<col style="width: 18%" />
<col style="width: 25%" />
<col style="width: 29%" />
<col style="width: 24%" />
<col style="width: 21%" />
<col style="width: 21%" />
</colgroup>
<thead>
<tr class="header">
<th><p>Trifolium eriocephalum</p></th>
<th><p>Crassula unknown</p></th>
<th><p>Castilleja attenuata</p></th>
<th><p>Triteleia laxa</p></th>
<th><p>Vicia villosa</p></th>
<th><p>Madia elegans</p></th>
<th><p>Geranium dissectum</p></th>
<th><p>Rumex pulcher</p></th>
<th><p>Vicia sativa¬†</p></th>
<th><p>Filago gallica</p></th>
<th><p>Avena barbata√</p></th>
<th><p>Annagalis arvensis√</p></th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td><ul>
<li><p>LDMC</p>
<ul>
<li><p>okay</p></li>
</ul></li>
<li><p>SLA</p>
<ul>
<li><p><strong>rep 2</strong></p></li>
<li><p>misenterred dry leaf mass√</p></li>
</ul></li>
<li><p>Coarse root specific length</p>
<ul>
<li><p>rep 2</p></li>
</ul></li>
<li><p>Fine root specific length</p>
<ul>
<li><p><strong>rep 3</strong></p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Coarse root specific length</p>
<ul>
<li><p><strong>rep 3</strong></p></li>
</ul></li>
<li><p>Fine root specific length</p>
<ul>
<li><p>rep 3</p></li>
</ul></li>
</ul>
<p>Checked data, dry weight is 0 even with microbalance, throwing these values off. √</p></td>
<td><ul>
<li><p>Coarse root specific length</p>
<ul>
<li><p>okay</p></li>
</ul></li>
<li><p>Fine root specific length</p>
<ul>
<li><p>okay</p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Coarse root specific length</p>
<ul>
<li><p>okay</p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Shoot dry biomass</p>
<ul>
<li><p>okay</p></li>
</ul></li>
<li><p>Coarse root length</p>
<ul>
<li><p>okay</p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Coarse root length</p>
<ul>
<li><p>okay</p></li>
<li><p>note two diff species with same name diff codes</p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Length</p>
<ul>
<li><p><strong>rep 4</strong></p></li>
</ul></li>
<li><p>Fine root length</p>
<ul>
<li><p>rep 4</p></li>
</ul></li>
</ul>
<p>Checked verbatim data, looks like the length is just really really long for this one.√</p></td>
<td><ul>
<li><p>Leaf Area</p>
<ul>
<li><p>okay</p></li>
</ul></li>
<li><p>Root dry biomass</p>
<ul>
<li><p>okay</p></li>
</ul></li>
<li><p>Total biomass</p>
<ul>
<li><p>okay</p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Shoot dry biomass</p>
<ul>
<li><p>okay</p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Fresh leaf mass</p>
<ul>
<li><p><strong>duplicates, off by a factor of 1000</strong></p></li>
</ul></li>
<li><p>SLA</p>
<ul>
<li><p>duplicates, off by a factor of 1000</p></li>
</ul></li>
</ul>
<p>Delete rows, duplicates but off by a factor of 100.</p>
<p>Cross referenced with raw 2021 greenhouse data√</p></td>
<td><ul>
<li><p>Height:</p>
<ul>
<li><p>171.700 value AVEBAR <strong>rep 2</strong></p></li>
</ul></li>
</ul></td>
<td><ul>
<li><p>Height</p>
<ul>
<li><p>13.0000 ANARp <strong>rep 4</strong></p></li>
</ul></li>
</ul></td>
</tr>
</tbody>
</table>

Cleaned data, new dataframe named: '**GreenhouseTraits_corrected.csv**' (annotations found in 'GreenhouseTraits_corrected_notes.xlsx'
